---
title: "Fitness_Complete_Fecundity_Analysis"
author: "Shaun Cross and Mark Stenglein"
date: "5/7/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

## Analyze Fitness Experiment #1 Fecundity span data

## Note: This is the complete analysis. Edits are made by Shaun Cross, but majority script designed by Mark's previous Rmd.

### Questions:
1. Do flies produce more offspring when infected with galbut virus?
2. Do flies change their rate/timing of offspring production when infected with galbut virus?
3. Do flies change the sex ratio of their offspring when infected with galbut virus?


## Experimental setup and data processing notes: 

**Expt'al setup:**
For strains 25192 and 25197: 
For each of these strains, there was a galbut-infected and a galbut-uninfected sub-strain
For each sub-strain: 5 replicate vials were setup with 5 M and 5 F adults (immediately after eclosion)
Every 3 days these original flies were transferred to new vials.
Two weeks later, the # of eclosed adult offspring in each vial was counted and the # of males and females

Shaun, Ali, and Tillie are recorded these offspring counts in this spreadsheet:
https://docs.google.com/spreadsheets/d/1GRPyScyD4YK8LGHbXkpnjkVacb3gZCfW0kjQn3PwqsU/edit#gid=329500492

Downloaded this complete data and put it into an Excel spreadsheet: Lifespan_complete_fecundity_tidy_data.xlsx
in a slightly cleaned up format.  

This analysis and the relevant files can be found in this google drive directory:
https://drive.google.com/drive/folders/1UVIeyWOw1Gw_1XB7BMZE5wkReud1KICY?usp=sharing

```{r warning=FALSE, message=FALSE}
# load libs
library(tidyverse)
library(readxl)
library(lubridate)
library(broom)
```


### Read the data in and do some pre-processing/summarizing

```{r}
#set working directory to where this file is saved
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# read file:
df <- read_excel("Lifespan_complete_fecundity_tidy_data.xlsx")

# rename BDSC strains IDs with DGRP names
df$strain <- str_replace(df$strain, "25192", "DGRP-399")
df$strain <- str_replace(df$strain, "25197", "DGRP-517")

# Calculate normalized offspring per adults in vial 
# This takes into account the fact that some parents begin to die over time
#
# This is not the perfect normalization to do since the surviving adults might be more males than females, but it ought to be reasonably close
#
# NOTE: this is configurable on the line below.
# I got essentially the same results when I didn't normalize.  There were fewer `
#
# keep track of pre-normalized values too
df <- df %>% mutate (non_normalized_male_offspring = male_offspring) 
df <- df %>% mutate (non_normalized_female_offspring = female_offspring)

normalize_offspring_counts <- 1
if (normalize_offspring_counts == TRUE) {
  df <- df %>% mutate (male_offspring = male_offspring / parents_in_vial)
  df <- df %>% mutate (female_offspring = female_offspring / parents_in_vial)
}

# calculate sex ratios
df <- df %>% mutate (sex_ratio = male_offspring / female_offspring)

# average replicates
df_avg <- df %>% group_by(strain, galbut, date_offspring_collected) %>% 
  summarize(avg_males = mean(male_offspring), std_males = sd(male_offspring), 
            avg_females = mean(female_offspring), std_females = sd(female_offspring),
            avg_parents = mean(parents_in_vial), std_parents = sd(parents_in_vial),
            avg_total = mean(emerged_offspring), std_total = sd(emerged_offspring)) 

# calculate cumulative sums of offspring per condition/replicate
df_sum <- df %>% group_by(strain, galbut, replicate) %>% mutate(cumulative_males = cumsum(male_offspring),
                                                                cumulative_females = cumsum(female_offspring))
# average cumulative replicates
df_sum_avg <- df_sum %>% group_by(strain, galbut, date_offspring_collected) %>% 
  summarize(avg_cumulative_males = mean(cumulative_males), std_males = sd(cumulative_males), 
            avg_cumulative_females = mean(cumulative_females), std_females = sd(cumulative_females)) 

# calculate un-normalized cumulative sums of offspring per condition/replicate
df_non_normalized_sum <- df %>% group_by(strain, galbut, replicate) %>% 
  mutate(cumulative_males = cumsum(non_normalized_male_offspring),
         cumulative_females = cumsum(non_normalized_female_offspring))

# nono-normalized average cumulative replicates
df_non_normalized_sum_avg <- df_non_normalized_sum %>% group_by(strain, galbut, date_offspring_collected) %>% 
  summarize(avg_cumulative_males = mean(cumulative_males), 
            std_males = sd(cumulative_males), 
            avg_cumulative_females = mean(cumulative_females), 
            std_females = sd(cumulative_females)) 
```

#### Make some plots to visualize results

```{r}
# plot sex ratios overall
ggplot(df) +
  geom_jitter(aes(x=galbut, y=sex_ratio, color=galbut), height=0, width=0.25) +
  facet_wrap(~strain) +
  theme_bw() + 
  theme(panel.grid=element_blank()) +
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  ggtitle("No difference in galbut infected/uninfected sex ratios") +
  xlab("galbut virus infection status") +
  ylab("male offspring / female offspring")

ggsave("offspring_sex_ratio.pdf", width = 5, height = 5, units = "in")
```

```{r}
# are these differences statistically significant? (no)
df %>% filter(is.finite(sex_ratio)) %>% group_by(strain) %>% do(tidy(t.test(sex_ratio ~ galbut, paired=FALSE, data=.)))
```


```{r}
# do sex ratios change over time?  (not obviously)
ggplot(df) +
  geom_jitter(aes(x=date_offspring_collected, y=sex_ratio, color=galbut), height=0, width=1)+
  facet_wrap(~strain) +
  theme_bw() + 
  theme(panel.grid=element_blank()) +
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  ggtitle("Sex ratios do not change over time in an obvious way") +
  xlab("galbut virus infection status") +
  ylab("male offspring / female offspring")
```

```{r}
# Plot #s of female offspring
ggplot(df) + 
  # geom_jitter(aes(x=date_offspring_collected,
  geom_jitter(aes(x=galbut,
                  y=female_offspring, 
                  color=galbut),
              height=0,
              width=0.1)+
  facet_wrap(~strain) +
  theme_bw() + 
  theme(panel.grid=element_blank()) +
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  ggtitle("Difference in # of female offspring between galbut infected and uninfected flies") +
  xlab("Galbut virus infection status") +
  ylab("Number female offspring per vial per surviving parent")
```

```{r}
# are these differences statistically significant? (no to both)
df %>% group_by(strain) %>% do(tidy(t.test(female_offspring ~ galbut, paired=FALSE, data=.)))
```

```{r}
# Plot #s of male offspring
ggplot(df) + 
  # geom_jitter(aes(x=date_offspring_collected,
  geom_jitter(aes(x=galbut,
                  y=male_offspring, 
                  color=galbut),
              height=0,
              width=0.1)+
  facet_wrap(~strain) +
  theme_bw() + 
  theme(panel.grid=element_blank()) +
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  ggtitle("Difference in # of male offspring between galbut infected and uninfected flies") +
  xlab("Galbut virus infection status") +
  ylab("Number male offspring per vial per surviving parent")
```

```{r}
# are these differences statistically significant? (399 no, 517 yes)
df %>% group_by(strain) %>% do(tidy(t.test(male_offspring ~ galbut, paired=FALSE, data=.)))
```

```{r}
#do total offspring vary based upon strain (outside of infection status)? Very sig!

View(df)
df %>% group_by(galbut) %>% do(tidy(t.test(emerged_offspring ~ strain, paired=FALSE, data=.)))

```

```{r}
#now let's plot the total number of offspring by strain

```

```{r}
# Plot cumulative #s of offspring
ggplot(df_sum) + 
  geom_line(aes(x=date_offspring_collected,
                  y=cumulative_males,
                  group=interaction(strain,galbut,replicate),
                  color=galbut)) + 
  geom_line(aes(x=date_offspring_collected,
                  y=cumulative_females,
                  group=interaction(strain,galbut,replicate),
                  color=galbut)) +
  facet_wrap(~strain) +
  theme_bw() + 
  theme(panel.grid=element_blank()) +
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  ggtitle("No difference in cumulative # of offspring between galbut infected and uninfected flies") +
  xlab("Date adult offspring counted") +
  ylab("Cumulative # of adult offspring per vial per surviving parent") 
```

```{r}
# Plot average cumulative #s of offspring
ggplot(df_sum_avg) + 
  geom_line(aes(x=date_offspring_collected,
                  y=avg_cumulative_males,
                  group=interaction(strain,galbut),
                  color=galbut),
            linetype=1) + 
  geom_errorbar(aes(x=date_offspring_collected, 
                    ymin = (avg_cumulative_males - std_males), 
                    ymax = (avg_cumulative_males + std_males), color=galbut), width = 10, size=0.2) +
  geom_line(aes(x=date_offspring_collected,
                  y=avg_cumulative_females,
                  group=interaction(strain,galbut),
                  color=galbut),  
            linetype=2) + 
  geom_errorbar(aes(x=date_offspring_collected, 
                    ymin = (avg_cumulative_females - std_females), 
                    ymax = (avg_cumulative_females + std_females), color=galbut), width = 1, size=0.2) +
  facet_wrap(~strain) +
  theme_bw() + 
  theme(panel.grid=element_blank()) +
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  # ggtitle("Slightly reduced cumulative # of offspring in galbut-infected 25192 flies\n25197 produce fewer female offspring (skewed sex ratio)\n(dotted line = female offspring)") +
  xlab("") +
  ylab("Cumulative # of adult offspring per surviving parent") 

#ggsave("cumulative_offspring_plot.pdf")
```

You can see that there is a lower # of cumulative offspring in the non-normalized cumulative offspring plot below for galbut-infected 21592.  This likely reflects a slightly reduced survivorship of the parents in galbut-infected 25192.

```{r}
# Plot NON-NORMALIZED average cumulative #s of offspring
ggplot(df_non_normalized_sum_avg) + 
  geom_line(aes(x=date_offspring_collected,
                  y=avg_cumulative_males,
                  group=interaction(strain,galbut),
                  color=galbut),
            linetype=1) + 
  geom_errorbar(aes(x=date_offspring_collected, 
                    ymin = (avg_cumulative_males - std_males), 
                    ymax = (avg_cumulative_males + std_males), color=galbut),  width = 10, size=0.2) +
  geom_line(aes(x=date_offspring_collected,
                  y=avg_cumulative_females,
                  group=interaction(strain,galbut),
                  color=galbut),  
            linetype=2) + 
  geom_errorbar(aes(x=date_offspring_collected, 
                    ymin = (avg_cumulative_females - std_females), 
                    ymax = (avg_cumulative_females + std_females), color=galbut),  width = 1, size=0.2) +
  facet_wrap(~strain, scales="free_y") +
  theme_bw() + 
  theme(panel.grid=element_blank()) +
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  ggtitle("Slightly reduced cumulative # of offspring in galbut-infected 25192 flies\n25197 produce fewer female offspring\n(dotted line = female offspring)\nnon-normalized version of above plot") +
  xlab("Date adult offspring counted") +
  ylab("Cumulative # of adult offspring per vial") 

#View(df_non_normalized_sum_avg)

```


#### Plot # of surviving parents 

```{r}
# Plot # of surviving parents 
ggplot(df_avg) + 
  geom_line(aes(x=date_offspring_collected,
                  y=avg_parents,
                  group=interaction(strain,galbut),
                  color=galbut),
            linetype=1) + 
  geom_errorbar(aes(x=date_offspring_collected, 
                    ymin = (avg_parents - std_parents), 
                    ymax = (avg_parents + std_parents), color=galbut),  width = 10, size=0.2) +
  facet_wrap(~strain) +
  theme_bw() + 
  theme(panel.grid=element_blank()) +
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  ggtitle("Slightly reduced survivorship of galbut-infected 25192 parents") +
  xlab("Date adult offspring counted") +
  ylab("Average # of parents in vial")
```
#### Change the dates into days since collected for better plotting
#### Will need to split df into 2 dataframes, then paste back together
```{r}
# Write function to convert dates to days
date_to_days <- function(x, start_date) {
  start_date <- start_date
  end <- ymd(x)
  elapsed.time <- start_date %--% end
  days_passed <- as.duration(elapsed.time) / ddays(1)
}

View(df_non_normalized_sum_avg_97)
df_non_normalized_sum_avg_97 <- df_non_normalized_sum_avg %>% filter(strain == "DGRP-517")
df_non_normalized_sum_avg_97 <- cbind(df_non_normalized_sum_avg_97, lapply(df_non_normalized_sum_avg_97[3], date_to_days, start_date = "2019-12-09"))
View(df_non_normalized_sum_avg_97)
names(df_non_normalized_sum_avg_97)[names(df_non_normalized_sum_avg_97) == "date_offspring_collected...8"] <- "days_offspring_collected"

df_non_normalized_sum_avg_92 <- df_non_normalized_sum_avg %>% filter(strain == "DGRP-399")
df_non_normalized_sum_avg_92 <- cbind(df_non_normalized_sum_avg_92, lapply(df_non_normalized_sum_avg_92[3], date_to_days, start_date = "2019-11-25"))
names(df_non_normalized_sum_avg_92)[names(df_non_normalized_sum_avg_92) == "date_offspring_collected...8"] <- "days_offspring_collected"

df_non_normalized_sum_avg_days <- bind_rows(df_non_normalized_sum_avg_92, df_non_normalized_sum_avg_97)

View(df_non_normalized_sum_avg_days)

```
# Plot this new days version
```{r}
# Plot NON-NORMALIZED average cumulative #s of offspring
ggplot(df_non_normalized_sum_avg_days) + 
  geom_line(aes(x=days_offspring_collected,
                  y=avg_cumulative_males,
                  group=interaction(strain,galbut),
                  color=galbut),
            linetype=1) + 
  geom_errorbar(aes(x=days_offspring_collected, 
                    ymin = (avg_cumulative_males - std_males), 
                    ymax = (avg_cumulative_males + std_males), color=galbut),  width = 1, size=0.2) +
  geom_line(aes(x=days_offspring_collected,
                  y=avg_cumulative_females,
                  group=interaction(strain,galbut),
                  color=galbut),  
            linetype=2) + 
 geom_errorbar(aes(x=days_offspring_collected, 
                    ymin = (avg_cumulative_females - std_females), 
                    ymax = (avg_cumulative_females + std_females), color=galbut),  width = 1, size=0.2) +
  facet_wrap(~strain, scales="free_y") +
  theme_bw() + 
  theme(panel.grid=element_blank()) +
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  scale_x_continuous(name="Days adult offspring counted", limits=c(14, 75)) +
  #scale_y_continuous(name="Cumulative # of adult offspring per vial", limits=c(0, 450))
  #ggtitle("Slightly reduced cumulative # of offspring in galbut-infected 25192 flies\n25197 produce fewer female offspring\n(dotted line = female offspring)\nnon-normalized version of above plot")
  #xlab("Date adult offspring counted") +
  ylab("Cumulative # of adult offspring per vial") 


ggsave("cummulative_offspring_days_axis.pdf")

```

#plotting total num of offspring by strain for supp fig
#this does not include galbut virus infected flies
#first filter out galbut virus positive flies, then plot

```{r}

#filter out galbut virus positive data
df_non_normalized_sum_avg_days_filt <- df_non_normalized_sum_avg_days %>% filter(galbut == FALSE)

ggplot(df_non_normalized_sum_avg_days_filt) + 
  geom_line(aes(x=days_offspring_collected,
                  y=avg_cumulative_males,
                  color=strain),
            linetype=1) + 
  geom_errorbar(aes(x=days_offspring_collected, 
                    ymin = (avg_cumulative_males - std_males), 
                    ymax = (avg_cumulative_males + std_males), color=strain),  width = 1, size=0.2) +
  geom_line(aes(x=days_offspring_collected,
                  y=avg_cumulative_females,
                  color=strain),  
            linetype=2) + 
 geom_errorbar(aes(x=days_offspring_collected, 
                    ymin = (avg_cumulative_females - std_females), 
                    ymax = (avg_cumulative_females + std_females), color=strain),  width = 1, size=0.2) +
  #facet_wrap(~strain, scales="free_y") +
  theme_bw() + 
  theme(panel.grid=element_blank()) +
  scale_color_manual(values = c("#009E73", "#CC79A7")) +
  scale_x_continuous(name="Days adult offspring counted", limits=c(14, 75)) +
  #scale_y_continuous(name="Cumulative # of adult offspring per vial", limits=c(0, 450))
  #ggtitle("Slightly reduced cumulative # of offspring in galbut-infected 25192 flies\n25197 produce fewer female offspring\n(dotted line = female offspring)\nnon-normalized version of above plot")
  #xlab("Date adult offspring counted") +
  ylab("Cumulative # of adult offspring per vial")
```