---
title: "Pathogen Challenge of Galbut-infected flies"
author: "Shaun T Cross"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

Galbut infected and uninfected DGRP 399 and 517 were challenged with various fly pathogens. This document looks to see if there is a significant difference in survival based on galbut virus infection status. This document makes use of the combined pathogen data sheet "pathogen_challenge_data_all.xlsx". The following loads in all appropriate libraries and sets the working directory where this file is saved.

```{r}
#set wd to where this file is saved
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

#load in libraries
library(tidyverse)
library(readxl)
library(lubridate)
library(ggpubr)
library(rstatix)
library(survival)
library(survminer)
library(cowplot)

```

## Drosophila C Virus (DCV) analysis

The first pathogen that was used was DCV. These steps evaluate the signficance of lifespan of galbut infected vs uninfected flies. First load in the data and change from strain stock number to DGRP number:

```{r}
df_DCV_surv <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "DCV_surv")
df_DCV_surv$strain <- str_replace(df_DCV_surv$strain, "25192", "DGRP-399")
df_DCV_surv$strain <- str_replace(df_DCV_surv$strain, "25197", "DGRP-517")
```

Chop up the data to make it best for analysis by sex and strain

```{r}
df_DCV_surv_399_M <- df_DCV_surv %>% filter(strain == "DGRP-399", sex == "male", DCV == TRUE)
df_DCV_surv_399_F <- df_DCV_surv %>% filter(strain == "DGRP-399", sex == "female", DCV == TRUE)

df_DCV_surv_517_M <- df_DCV_surv %>% filter(strain == "DGRP-517", sex == "male", DCV == TRUE)
df_DCV_surv_517_F <- df_DCV_surv %>% filter(strain == "DGRP-517", sex == "female", DCV == TRUE)

```

Run the survival analysis. See: https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html for an excellent introduction to survival analysis in R. Also see: https://rviews.rstudio.com/2017/09/25/survival-analysis-with-r/

Any flies that either did not die by the experimental period or via unnatural causes (e.g. stuck in media or from microinjection) is censored.

```{r}
sd_DCV_399_M <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_DCV_surv_399_M)
sd_DCV_399_F <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_DCV_surv_399_F)

sd_DCV_517_M <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_DCV_surv_517_M)
sd_DCV_517_F <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_DCV_surv_517_F)

#survival analysis
1 - pchisq(sd_DCV_399_M$chisq, length(sd_DCV_399_M$n) - 1) # p = 0.224
1 - pchisq(sd_DCV_399_F$chisq, length(sd_DCV_399_F$n) - 1) # p = 0.341

1 - pchisq(sd_DCV_517_M$chisq, length(sd_DCV_517_M$n) - 1) # p = 0.271
1 - pchisq(sd_DCV_517_F$chisq, length(sd_DCV_517_F$n) - 1) # p = 0.0279

```
Now we are going to plot the survival curve. Givent the small time frame, the plotting function found in the survival package looks clunky. Choosing to plot with ggplot2. To do this, all censored flies were removed from the total to give number of valid flies for plotting. Then plotted as found below:

```{r}

df_tidy_DCV <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "DCV_tidy")

df_tidy_DCV$strain <- str_replace(df_tidy_DCV$strain, "25197", "DGRP-517")
df_tidy_DCV$strain <- str_replace(df_tidy_DCV$strain, "25192", "DGRP-399")

# average replicates
df_tidy_DCV_avg <- df_tidy_DCV %>% group_by(strain, galbut, sex, day) %>% summarize(avg_alive = mean(percent_alive), std_alive = sd(percent_alive)) 

# plot it
p_DCV <- ggplot(df_tidy_DCV_avg, aes(x=day, 
                                         y=avg_alive, 
                                         group=interaction(strain, galbut, sex)))  + 
  geom_point(size=0.5) +
  geom_line(aes(color=galbut)) +
  geom_errorbar(aes(ymin = (avg_alive - std_alive), ymax = (avg_alive + std_alive)), width = 0.2, color="grey50", size=0.2) +
  facet_grid(sex~strain) + 
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  theme_bw() + 
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5, size = 9),
        legend.position = "none",
        axis.text = element_text(size = 7),
        axis.title = element_text(size = 8),
        strip.text.x = element_text(size = 7.5),
        strip.text.y = element_text(size = 7.5)) +
  scale_x_continuous(breaks=seq(0,14,2)) +
  ggtitle("Drosophila C Virus Injected Flies") +
  xlab("Days post injection")  +
  ylab("Mean percent flies alive per vial")

p_DCV

```
NOTE: When I plot these figures, I have removed the legend which is helpful for fitting all plots together in the final figure. To keep the legend, just remove the following line of code:

```{r}
# legend.position = "none"
```


## Pseudomonas aeruginosa oral challenge analysis

The next pathogen that was used was P. aeurginosa (PA) via oral challenge. These steps evaluate the signficance of lifespan of galbut infected vs uninfected flies. All same analysis as above. First load in the data and change from strain stock number to DGRP number:

```{r}
df_PA_surv_feed <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "PA_feed_surv")

df_PA_surv_feed$strain <- str_replace(df_PA_surv_feed$strain, "25192", "DGRP-399")
df_PA_surv_feed$strain <- str_replace(df_PA_surv_feed$strain, "25197", "DGRP-517")
```
Chop up the data for sex and strain.

```{r}
df_PA_surv_feed_399_M <- df_PA_surv_feed %>% filter(strain == "DGRP-399", sex == "male", pseudomonas == TRUE)
df_PA_surv_feed_399_F <- df_PA_surv_feed %>% filter(strain == "DGRP-399", sex == "female", pseudomonas == TRUE)

df_PA_surv_feed_517_M <- df_PA_surv_feed %>% filter(strain == "DGRP-517", sex == "male", pseudomonas == TRUE)
df_PA_surv_feed_517_F <- df_PA_surv_feed %>% filter(strain == "DGRP-517", sex == "female", pseudomonas == TRUE)
```
Survival analysis

```{r}
sd_PA_surv_feed_399_M <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_PA_surv_feed_399_M)
sd_PA_surv_feed_399_F <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_PA_surv_feed_399_F)

sd_PA_surv_feed_517_M <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_PA_surv_feed_517_M)
sd_PA_surv_feed_517_F <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_PA_surv_feed_517_F)

#survival analysis
1 - pchisq(sd_PA_surv_feed_399_M$chisq, length(sd_PA_surv_feed_399_M$n) - 1) # p = 0.3001
1 - pchisq(sd_PA_surv_feed_399_F$chisq, length(sd_PA_surv_feed_399_M$n) - 1) # p = 4.5e-6

1 - pchisq(sd_PA_surv_feed_517_M$chisq, length(sd_PA_surv_feed_517_M$n) - 1) # p = 0.42
1 - pchisq(sd_PA_surv_feed_517_F$chisq, length(sd_PA_surv_feed_517_F$n) - 1) # p = 0.729

```
Using tidy data, plot survival curves

```{r}
df_tidy_pseudomonas_feed <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "PA_feed_tidy")

df_tidy_pseudomonas_feed$strain <- str_replace(df_tidy_pseudomonas_feed$strain, "25197", "DGRP-517")
df_tidy_pseudomonas_feed$strain <- str_replace(df_tidy_pseudomonas_feed$strain, "25192", "DGRP-399")

# average replicates
df_tidy_pseudomonas_feed_avg <- df_tidy_pseudomonas_feed %>% group_by(strain, galbut, sex, day) %>% summarize(avg_alive = mean(percent_alive), std_alive = sd(percent_alive)) 



p_PA_feed <- ggplot(df_tidy_pseudomonas_feed_avg, aes(x=day, 
                                                y=avg_alive, 
                                                group=interaction(strain, galbut, sex)))  + 
  geom_point(size=0.5) +
  geom_line(aes(color=galbut)) +
  geom_errorbar(aes(ymin = (avg_alive - std_alive), ymax = (avg_alive + std_alive)), width = 0.2, color="grey50", size=0.2) +
  facet_grid(sex~strain) + 
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  theme_bw() + 
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5, size = 9),
        legend.position = "none",
        axis.text = element_text(size = 7),
        axis.title = element_text(size = 8),
        strip.text.x = element_text(size = 7.5),
        strip.text.y = element_text(size = 7.5)) +
  coord_cartesian(ylim=c(0,100)) + #have to use this to rescale as other scaling makes the error bars disappear
  scale_x_continuous(breaks=seq(0,12,2)) +
  ggtitle("Pseudomonas aeruginosa Orally Challenged Flies") +
  xlab("Days since oral challenge began")  +
  ylab("Mean percent flies alive per vial")

p_PA_feed
```
## Pseudomonas aeruginosa inject challenge analysis

Oral inoculation has limitations in reproducibility as it is highly dependent on how much the flies eat. Injection allows for an exact dosage. Did PA injection next.

```{r}
df_PA_surv_inject <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "PA_inject_surv")

df_PA_surv_inject$strain <- str_replace(df_PA_surv_inject$strain, "25197", "DGRP-517")
df_PA_surv_inject$strain <- str_replace(df_PA_surv_inject$strain, "25192", "DGRP-399")
```

Split data for sex and strain analysis

```{r}
df_PA_surv_inject_399_M <- df_PA_surv_inject %>% filter(strain == "DGRP-399", sex == "male", pseudomonas == TRUE)
df_PA_surv_inject_399_F <- df_PA_surv_inject %>% filter(strain == "DGRP-399", sex == "female", pseudomonas == TRUE)

df_PA_surv_inject_517_M <- df_PA_surv_inject %>% filter(strain == "DGRP-517", sex == "male", pseudomonas == TRUE)
df_PA_surv_inject_517_F <- df_PA_surv_inject %>% filter(strain == "DGRP-517", sex == "female", pseudomonas == TRUE)

```
Run the survival analysis. Same as above, flies that died unnaturally are censored.

```{r}
sd_PA_surv_inject_399_M <- survdiff(Surv(hours_survived, !censor) ~ galbut, data = df_PA_surv_inject_399_M)
sd_PA_surv_inject_399_F <- survdiff(Surv(hours_survived, !censor) ~ galbut, data = df_PA_surv_inject_399_F)

sd_PA_surv_inject_517_M <- survdiff(Surv(hours_survived, !censor) ~ galbut, data = df_PA_surv_inject_517_M)
sd_PA_surv_inject_517_F <- survdiff(Surv(hours_survived, !censor) ~ galbut, data = df_PA_surv_inject_517_F)

#survival analysis
1 - pchisq(sd_PA_surv_inject_399_M$chisq, length(sd_PA_surv_inject_399_M$n) - 1) # p = 0.8907
1 - pchisq(sd_PA_surv_inject_399_F$chisq, length(sd_PA_surv_inject_399_F$n) - 1) # p = 0.1362

1 - pchisq(sd_PA_surv_inject_517_M$chisq, length(sd_PA_surv_inject_517_M$n) - 1) # p = 0.2183
1 - pchisq(sd_PA_surv_inject_517_F$chisq, length(sd_PA_surv_inject_517_F$n) - 1) # p = 0.2112
```
Now plot the survival curves.

```{r}
df_tidy_pseudomonas_inject <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "PA_inject_tidy")

df_tidy_pseudomonas_inject$strain <- str_replace(df_tidy_pseudomonas_inject$strain, "25197", "DGRP-517")
df_tidy_pseudomonas_inject$strain <- str_replace(df_tidy_pseudomonas_inject$strain, "25192", "DGRP-399")

# average replicates
df_tidy_pseudomonas_inject_avg <- df_tidy_pseudomonas_inject %>% group_by(strain, galbut, sex, hours) %>% summarize(avg_alive = mean(percent_alive), std_alive = sd(percent_alive)) 

# plot it
p_PA_inject <- ggplot(df_tidy_pseudomonas_inject_avg, aes(x=factor(hours), 
                              y=avg_alive, 
                              group=interaction(strain, galbut, sex)))  + 
  geom_point(size=0.5) +
  geom_line(aes(color=galbut)) +
  geom_errorbar(aes(ymin = (avg_alive - std_alive), ymax = (avg_alive + std_alive)), width = 0.2, color="grey50", size=0.2) +
  facet_grid(sex~strain) + 
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  theme_bw() + 
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5, size = 9),
        legend.position = "none",
        axis.text = element_text(size = 7),
        axis.text.x = element_text(size = 6),
        axis.title = element_text(size = 8),
        strip.text.x = element_text(size = 7.5),
        strip.text.y = element_text(size = 7.5)) +
  ggtitle("Pseudomonas aeruginosa Injected Flies") +
  xlab("Hours post injection")  +
  ylab("Mean percent flies alive per vial")

p_PA_inject
```

## Staphylococcus aureus inject challenge analysis
P. aeruginosa is a Gram neg bacteria. Moved onto a Gram positive pathogen challenge with S. aureus (SA).

```{r}
df_surv_staph <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "SA_surv")

df_surv_staph$strain <- str_replace(df_surv_staph$strain, "25197", "DGRP-517")
df_surv_staph$strain <- str_replace(df_surv_staph$strain, "25192", "DGRP-399")

```

Get data ready again and then run analysis.

```{r}
df_surv_399_M_SA <- df_surv_staph %>% filter(strain == "DGRP-399", sex == "male", staphylococcus == TRUE)
df_surv_399_F_SA <- df_surv_staph %>% filter(strain == "DGRP-399", sex == "female", staphylococcus == TRUE)


df_surv_517_M_SA <- df_surv_staph %>% filter(strain == "DGRP-517", sex == "male", staphylococcus == TRUE)
df_surv_517_F_SA <- df_surv_staph %>% filter(strain == "DGRP-517", sex == "female", staphylococcus == TRUE)



sd_399_M_SA <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_399_M_SA)
sd_399_F_SA <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_399_F_SA)

sd_517_M_SA <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_517_M_SA)
sd_517_F_SA <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_517_F_SA)

#survival analysis
1 - pchisq(sd_399_M_SA$chisq, length(sd_399_M_SA$n) - 1) # p = 0.0207
1 - pchisq(sd_399_F_SA$chisq, length(sd_399_F_SA$n) - 1) # p = 0.0549

1 - pchisq(sd_517_M_SA$chisq, length(sd_517_M_SA$n) - 1) # p = 0.0509
1 - pchisq(sd_517_F_SA$chisq, length(sd_517_F_SA$n) - 1) # p = 0.00068
```
Plot survival curves using tidy data

```{r}
df_tidy_staph <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "SA_tidy")

df_tidy_staph$strain <- str_replace(df_tidy_staph$strain, "25197", "DGRP-517")
df_tidy_staph$strain <- str_replace(df_tidy_staph$strain, "25192", "DGRP-399")

# average replicates
df_tidy_staph_avg <- df_tidy_staph %>% group_by(strain, galbut, sex, day) %>% summarize(avg_alive = mean(percent_alive), std_alive = sd(percent_alive)) 

# plot it
p_staph <- ggplot(df_tidy_staph_avg, aes(x=day, 
                   y=avg_alive, 
                   group=interaction(strain, galbut, sex)))  + 
  geom_point(size=0.5) +
  geom_line(aes(color=galbut)) +
  geom_errorbar(aes(ymin = (avg_alive - std_alive), ymax = (avg_alive + std_alive)), width = 0.2, color="grey50", size=0.2) +
  facet_grid(sex~strain) + 
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  theme_bw() + 
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5, size = 9),
        legend.position = "none",
        axis.text = element_text(size = 7),
        axis.title = element_text(size = 8),
        strip.text.x = element_text(size = 7.5),
        strip.text.y = element_text(size = 7.5)) +
  ggtitle("Staphylococcus aureus Injected Flies") +
  xlab("Days post injection")  +
  ylab("Mean percent flies alive per vial")

p_staph
```
## Candida albicans inject challenge analysis

Finally, we rounded out our pathogen experiments by challenging with a fungal pathogen, Candida albicans (CA).

```{r}
df_surv_CA <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "CA_surv")

df_surv_CA$strain <- str_replace(df_surv_CA$strain, "25197", "DGRP-517")
df_surv_CA$strain <- str_replace(df_surv_CA$strain, "25192", "DGRP-399")

```

Chop up the data and run the analysis.

```{r}
df_surv_399_M_CA <- df_surv_CA %>% filter(strain == "DGRP-399", sex == "male", candida == TRUE)
df_surv_399_F_CA <- df_surv_CA %>% filter(strain == "DGRP-399", sex == "female", candida == TRUE)


df_surv_517_M_CA <- df_surv_CA %>% filter(strain == "DGRP-517", sex == "male", candida == TRUE)
df_surv_517_F_CA <- df_surv_CA %>% filter(strain == "DGRP-517", sex == "female", candida == TRUE)

#run the analysis
sd_399_M_CA <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_399_M_CA)
sd_399_F_CA <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_399_F_CA)

sd_517_M_CA <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_517_M_CA)
sd_517_F_CA <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_517_F_CA)

#survival analysis
1 - pchisq(sd_399_M_CA$chisq, length(sd_399_M_CA$n) - 1) # p = 6.5108e-6
1 - pchisq(sd_399_F_CA$chisq, length(sd_399_F_CA$n) - 1) # p = 3.4735e-5

1 - pchisq(sd_517_M_CA$chisq, length(sd_517_M_CA$n) - 1) # p = 0.1238
1 - pchisq(sd_517_F_CA$chisq, length(sd_517_F_CA$n) - 1) # p = 0.7786

```
Plot the survival curves using tidy data

```{r}
df_tidy_CA <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "CA_tidy")

df_tidy_CA$strain <- str_replace(df_tidy_CA$strain, "25197", "DGRP-517")
df_tidy_CA$strain <- str_replace(df_tidy_CA$strain, "25192", "DGRP-399")

# average replicates
df_tidy_CA_avg <- df_tidy_CA %>% group_by(strain, galbut, sex, day) %>% summarize(avg_alive = mean(percent_alive), std_alive = sd(percent_alive)) 

# plot it
p_CA <- ggplot(df_tidy_CA_avg, aes(x=day, 
                                     y=avg_alive, 
                                     group=interaction(strain, galbut, sex)))  + 
  geom_point(size=0.5) +
  geom_line(aes(color=galbut)) +
  geom_errorbar(aes(ymin = (avg_alive - std_alive), ymax = (avg_alive + std_alive)), width = 0.2, color="grey50", size=0.2) +
  facet_grid(sex~strain) + 
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  theme_bw() + 
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5, size = 9),
        legend.position = "none",
        axis.text = element_text(size = 7),
        axis.title = element_text(size = 8),
        strip.text.x = element_text(size = 7.5),
        strip.text.y = element_text(size = 7.5)) +
  ggtitle("Candida albicans Injected Flies") +
  coord_cartesian(ylim=c(0,100)) + #have to use this to rescale as other scaling makes the error bars disappear
  xlab("Days post injection")  +
  ylab("Mean percent flies alive per vial")

p_CA

```
## Control analysis
We ran controls alongside these infections. We have 3 control groups. Those that were injected with PBS and held at 25deg C (DCV, PA inject, Staph inject), those that were injected with PBS and held at 30deg C (CA), and those that were fed sucrose (PA oral challenge). The following analyzes the data.

First PBS 25deg analysis
```{r}
df_surv_PBS_25 <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "PBS_25_surv")

df_surv_PBS_25$strain <- str_replace(df_surv_PBS_25$strain, "25197", "DGRP-517")
df_surv_PBS_25$strain <- str_replace(df_surv_PBS_25$strain, "25192", "DGRP-399")
```

chop the data and do the analysis

```{r}
df_surv_399_M_PBS_25 <- df_surv_PBS_25 %>% filter(strain == "DGRP-399", sex == "male", PBS == TRUE)
df_surv_399_F_PBS_25 <- df_surv_PBS_25 %>% filter(strain == "DGRP-399", sex == "female", PBS == TRUE)


df_surv_517_M_PBS_25 <- df_surv_PBS_25 %>% filter(strain == "DGRP-517", sex == "male", PBS == TRUE)
df_surv_517_F_PBS_25 <- df_surv_PBS_25 %>% filter(strain == "DGRP-517", sex == "female", PBS == TRUE)

#run the analysis
sd_surv_399_M_PBS_25 <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_399_M_PBS_25)
sd_surv_399_F_PBS_25 <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_399_F_PBS_25)

sd_surv_517_M_PBS_25 <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_517_M_PBS_25)
sd_surv_517_F_PBS_25 <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_517_F_PBS_25)

#survival analysis
1 - pchisq(sd_surv_399_M_PBS_25$chisq, length(sd_surv_399_M_PBS_25$n) - 1) # p = 0.461
1 - pchisq(sd_surv_399_F_PBS_25$chisq, length(sd_surv_399_F_PBS_25$n) - 1) # p = 0.398

1 - pchisq(sd_surv_517_M_PBS_25$chisq, length(sd_surv_517_M_PBS_25$n) - 1) # p = 0.0547
1 - pchisq(sd_surv_517_F_PBS_25$chisq, length(sd_surv_517_F_PBS_25$n) - 1) # p = 0.107
```

Then PBS 30deg analysis
```{r}
df_surv_PBS_30 <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "PBS_30_surv")

df_surv_PBS_30$strain <- str_replace(df_surv_PBS_30$strain, "25197", "DGRP-517")
df_surv_PBS_30$strain <- str_replace(df_surv_PBS_30$strain, "25192", "DGRP-399")
```

chop the data and do the analysis

```{r}
df_surv_399_M_PBS_30 <- df_surv_PBS_30 %>% filter(strain == "DGRP-399", sex == "male", candida == FALSE)
df_surv_399_F_PBS_30 <- df_surv_PBS_30 %>% filter(strain == "DGRP-399", sex == "female", candida == FALSE)


df_surv_517_M_PBS_30 <- df_surv_PBS_30 %>% filter(strain == "DGRP-517", sex == "male", candida == FALSE)
df_surv_517_F_PBS_30 <- df_surv_PBS_30 %>% filter(strain == "DGRP-517", sex == "female", candida == FALSE)

#run the analysis
sd_surv_399_M_PBS_30 <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_399_M_PBS_30)
sd_surv_399_F_PBS_30 <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_399_F_PBS_30)

sd_surv_517_M_PBS_30 <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_517_M_PBS_30)
sd_surv_517_F_PBS_30 <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_517_F_PBS_30)

#survival analysis
1 - pchisq(sd_surv_399_M_PBS_30$chisq, length(sd_surv_399_M_PBS_30$n) - 1) # p = 0.971
1 - pchisq(sd_surv_399_F_PBS_30$chisq, length(sd_surv_399_F_PBS_30$n) - 1) # p = 0.777

1 - pchisq(sd_surv_517_M_PBS_30$chisq, length(sd_surv_517_M_PBS_30$n) - 1) # p = 0.0738
1 - pchisq(sd_surv_517_F_PBS_30$chisq, length(sd_surv_517_F_PBS_30$n) - 1) # p = 0.622

```

Finally sucrose analysis
```{r}
df_surv_sucrose <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "sucrose_surv")

df_surv_sucrose$strain <- str_replace(df_surv_sucrose$strain, "25197", "DGRP-517")
df_surv_sucrose$strain <- str_replace(df_surv_sucrose$strain, "25192", "DGRP-399")
```

chop the data and do the analysis

```{r}
df_surv_399_M_sucrose <- df_surv_sucrose %>% filter(strain == "DGRP-399", sex == "male", pseudomonas == FALSE)
df_surv_399_F_sucrose <- df_surv_sucrose %>% filter(strain == "DGRP-399", sex == "female", pseudomonas == FALSE)


df_surv_517_M_sucrose <- df_surv_sucrose %>% filter(strain == "DGRP-517", sex == "male", pseudomonas == FALSE)
df_surv_517_F_sucrose <- df_surv_sucrose %>% filter(strain == "DGRP-517", sex == "female", pseudomonas == FALSE)

#run the analysis
sd_surv_399_M_sucrose <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_399_M_sucrose)
sd_surv_399_F_sucrose <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_399_F_sucrose)

sd_surv_517_M_sucrose <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_517_M_sucrose)
sd_surv_517_F_sucrose <- survdiff(Surv(days_survived, !censor) ~ galbut, data = df_surv_517_F_sucrose)

#survival analysis
1 - pchisq(sd_surv_399_M_sucrose$chisq, length(sd_surv_399_M_sucrose$n) - 1) # p = 0.0116
1 - pchisq(sd_surv_399_F_sucrose$chisq, length(sd_surv_399_F_sucrose$n) - 1) # p = 0.0647

1 - pchisq(sd_surv_517_M_sucrose$chisq, length(sd_surv_517_M_sucrose$n) - 1) # p = 0.00292
1 - pchisq(sd_surv_517_F_sucrose$chisq, length(sd_surv_517_F_sucrose$n) - 1) # p = 0.184

```
Only sucrose feeding controls had significant differences. This means there are likely some confidence issues in what we measure in the experimental groups as well.

Plotting all controls on the same plot.

```{r}
df_tidy_controls <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "controls_combined_tidy")

df_tidy_controls$strain <- str_replace(df_tidy_controls$strain, "25197", "DGRP-517")
df_tidy_controls$strain <- str_replace(df_tidy_controls$strain, "25192", "DGRP-399")

# average replicates
df_tidy_controls_avg <- df_tidy_controls %>% group_by(strain, galbut, sex, day, cont_type) %>% summarize(avg_alive = mean(percent_alive), std_alive = sd(percent_alive)) 

# plot it
p_controls <- ggplot(df_tidy_controls_avg, aes(x=day, 
                                     y=avg_alive, 
                                     group=interaction(strain, galbut, sex, cont_type)))  + 
  geom_point(size=0.5) +
  geom_line(aes(color=galbut, linetype = cont_type)) +
  geom_errorbar(aes(ymin = (avg_alive - std_alive), ymax = (avg_alive + std_alive)), width = 0.2, color="grey50", size=0.2) +
  facet_grid(sex~strain) + 
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  theme_bw() + 
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5, size = 9),
        legend.position = "none",
        axis.text = element_text(size = 7),
        axis.title = element_text(size = 8),
        strip.text.x = element_text(size = 7.5),
        strip.text.y = element_text(size = 7.5)) +
  ggtitle("Control Challenged Flies") +
  coord_cartesian(ylim=c(0,100)) + #have to use this to rescale as other scaling makes the error bars disappear
  xlab("Days post injection")  +
  ylab("Mean percent flies alive per vial")

p_controls
```
Looks congested though, so will split each one into a separate plot.

First do PBS 25deg C controls
```{r}
df_tidy_PBS_25 <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "PBS_25_tidy")

df_tidy_PBS_25$strain <- str_replace(df_tidy_PBS_25$strain, "25197", "DGRP-517")
df_tidy_PBS_25$strain <- str_replace(df_tidy_PBS_25$strain, "25192", "DGRP-399")

# average replicates
df_tidy_PBS_25_avg <- df_tidy_PBS_25 %>% group_by(strain, galbut, sex, day) %>% summarize(avg_alive = mean(percent_alive), std_alive = sd(percent_alive)) 

# plot it
p_PBS_25 <- ggplot(df_tidy_PBS_25_avg, aes(x=day, 
                                     y=avg_alive, 
                                     group=interaction(strain, galbut, sex)))  + 
  geom_point(size=0.5) +
  geom_line(aes(color=galbut)) +
  geom_errorbar(aes(ymin = (avg_alive - std_alive), ymax = (avg_alive + std_alive)), width = 0.2, color="grey50", size=0.2) +
  facet_grid(sex~strain) + 
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  theme_bw() + 
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5, size = 9),
        legend.position = "none",
        axis.text = element_text(size = 7),
        axis.title = element_text(size = 8),
        strip.text.x = element_text(size = 7.5),
        strip.text.y = element_text(size = 7.5)) +
  scale_x_continuous(breaks=seq(0,14,2)) +
  ggtitle("PBS 25deg Control Injected Flies") +
  coord_cartesian(ylim=c(0,100)) + #have to use this to rescale as other scaling makes the error bars disappear
  xlab("Days post injection")  +
  ylab("Mean percent flies alive per vial")

p_PBS_25
```

Now for 30deg C control flies

```{r}
df_tidy_PBS_30 <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "PBS_30_tidy")

df_tidy_PBS_30$strain <- str_replace(df_tidy_PBS_30$strain, "25197", "DGRP-517")
df_tidy_PBS_30$strain <- str_replace(df_tidy_PBS_30$strain, "25192", "DGRP-399")

# average replicates
df_tidy_PBS_30_avg <- df_tidy_PBS_30 %>% group_by(strain, galbut, sex, day) %>% summarize(avg_alive = mean(percent_alive), std_alive = sd(percent_alive)) 

# plot it
p_PBS_30 <- ggplot(df_tidy_PBS_30_avg, aes(x=day, 
                                     y=avg_alive, 
                                     group=interaction(strain, galbut, sex)))  + 
  geom_point(size=0.5) +
  geom_line(aes(color=galbut)) +
  geom_errorbar(aes(ymin = (avg_alive - std_alive), ymax = (avg_alive + std_alive)), width = 0.2, color="grey50", size=0.2) +
  facet_grid(sex~strain) + 
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  theme_bw() + 
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5, size = 9),
        legend.position = "none",
        axis.text = element_text(size = 7),
        axis.title = element_text(size = 8),
        strip.text.x = element_text(size = 7.5),
        strip.text.y = element_text(size = 7.5)) +
  ggtitle("PBS 30deg Control Injected Flies") +
  coord_cartesian(ylim=c(0,100)) + #have to use this to rescale as other scaling makes the error bars disappear
  xlab("Days post injection")  +
  ylab("Mean percent flies alive per vial")

p_PBS_30
```

Finally doing it for sucrose controls

```{r}
df_tidy_sucrose <- read_excel("pathogen_challenge_data_all.xlsx", sheet = "sucrose_tidy")

df_tidy_sucrose$strain <- str_replace(df_tidy_sucrose$strain, "25197", "DGRP-517")
df_tidy_sucrose$strain <- str_replace(df_tidy_sucrose$strain, "25192", "DGRP-399")

# average replicates
df_tidy_sucrose_avg <- df_tidy_sucrose %>% group_by(strain, galbut, sex, day) %>% summarize(avg_alive = mean(percent_alive), std_alive = sd(percent_alive)) 

# plot it
p_sucrose <- ggplot(df_tidy_sucrose_avg, aes(x=day, 
                                     y=avg_alive, 
                                     group=interaction(strain, galbut, sex)))  + 
  geom_point(size=0.5) +
  geom_line(aes(color=galbut)) +
  geom_errorbar(aes(ymin = (avg_alive - std_alive), ymax = (avg_alive + std_alive)), width = 0.2, color="grey50", size=0.2) +
  facet_grid(sex~strain) + 
  scale_color_manual(values = c("#56B4E9", "#D55E00")) +
  theme_bw() + 
  theme(panel.grid=element_blank(),
        plot.title = element_text(hjust = 0.5, size = 9),
        legend.position = "none",
        axis.text = element_text(size = 7),
        axis.title = element_text(size = 8),
        strip.text.x = element_text(size = 7.5),
        strip.text.y = element_text(size = 7.5)) +
  ggtitle("Sucrose Control Challenged Flies") +
  coord_cartesian(ylim=c(0,100)) + #have to use this to rescale as other scaling makes the error bars disappear
  scale_x_continuous(breaks=seq(0,12,2)) +
  xlab("Days post injection")  +
  ylab("Mean percent flies alive per vial")

p_sucrose
```

Now combine all the plots into one for the figure.

```{r}
plot_grid(p_DCV, p_PA_feed, p_PA_inject, p_staph, p_CA, p_PBS_25, p_PBS_30, p_sucrose,
          labels = "AUTO", label_size = 11,
          ncol = 2)
ggsave("pathogen_challenge_plot_v2.pdf", width = 6.5, height = 8.5, units = "in")

```
